### 2015-07-13 manuscript/defs2.tex
Updating defs2

\begin{definition}
A \underline{repeat element family discriptor} (refd) is a sequence over
$\Sigma=\{A,C,G,T,*\}$, where $*$ denotes unspecified bases. 
\end{definition}

{\bf Note:} We will refer to sub-refds as consistent with the concept
of a substring, and denote then as consistent with Python.  (That is,
if $r =  AAAC*CGG'$, then the sub-refd $r_{3:6}$, or $r[3:6]$, is the refd
$C*C$).  Generally, we will define all standard string operators on
refds unless otherwise noted.  (e.g. $\cdot$ will be the concatenation
operator, $|r|$ will denote total length, etc...)

@@

\begin{definition}
Let $r$ and $r'$ be two $l$-mers.
\begin{itemize}
\item $r$ and $r'$ are \underline{consistent} if,
  for any position $i$ in which they differ, one of them contains a
  $*$.  ($AAC*TT$ and $AA*GTT$ are consistent; $AAC*TT$ and $AAGGTT$
  are not.)
\item We will say $r$ \underline{generalizes} $r'$ if, wherever they differ,
  $r$ has a $*$.  Denote this as $r \vdash r'$.  ($AAC*TT \vdash AACGTT$, 
  $AAC*TT \not\vdash AA*CTT$.)
\item The \underline{merge} of $r$ and $r'$, denoted $r \oplus r'$, is
  the $l$-mer such that $s_i = r_i$ when $r_i=r'_i$,
  and $*$ otherwise. ($AAC*TT \oplus AA*GTT = AA**TT$.) 
\item The \underline{tight merge} of $r$ and $r'$, denoted $r \otimes r'$,
  is $r \oplus r'$ if $r$ and $r'$ are consistent, and undefined
  otherwise. ($AAC*TT \otimes AA*GTT = AA**GT$, but $AAC*TT \otimes AAAA*CGT$ is undefined.)
\item The \underline{compression} of consistent refds $r$ and $r'$, denoted $r\ominus r'$, 
  is the string $s$ where $s_i$ is: $*$ if both $r_i$ and $r'_i$ are
  $*$, other the non-$*$ character at $r_i$ or $r'_i$.  If $r$ is and
  $r'$ are not consistent, $r \ominus r'$ is undefined.  ($AA**C
  \ominus A**CC = AA*CC$.
\end{itemize}
\end{definition}

@@

\begin{definition}
Let $x$ and $y$ be REFDs. Define $x \circ y$ as follows.  Let $i$ be
the largest $i$ such that $x_{|x|-i:}$ and $y_{:i}$ are consistent, let
$b = x_{|x|-i:} \ominus y_{:i}$, let $a = x_{:|x|-i}$ and $c = y_{i:}$.  
$x \circ y = a \cdot b \cdot c$.
\end{definition}
{\bf Examples:}
\begin{itemize}
\item $AAACCC \circ C*CTTT = AACCC$.
\item $AAACCC \circ CTCTTT = AAATCCTCTTT$.
\item $AAACC* \circ CC*TTT = AAACC*TTT$.
\end{itemize}

@@

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, The seeded
$l$-mer, denoted $\sigma_s(r)$, is created by removing all bases at
positions corresponding to 0s.
\end{definition}
Example: $\sigma_{11011}(AACGG) = AAGG$.

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
generlized $l$-mer (w.r.t. $s$), or $\gamma_s(r)$, is the refd created by replacing
all basses corresponding to 0s with * characters.
\end{definition}
Example: $\sigma_{110101}(AACGTG) = AA*G*G$.

@@

\begin{definition} 
A spaced seed $s$ of length $l$ {\underline covers} and refd $r$ is, for every $j$,
there is an $i \leq j$ such that $s$ is consistent with the sub-refd
$r_{i:i+|l|}$.
\end{definition}
{\bf Examples:} Let $s = 11011$.
\begin{itemize}
\item $r = AAAAA*AA$. $s$ covers $r$ as we can place $s$ $i=0$ and
  $i=3$ and, between them, cover every character in $r$ without
  conclict between a 1 and a $*$.
\item $r = AAA*AA$.  $s$ does not cover $r$.  $s$ only hist at $i=1$,
  so there is no way to cover base 0.
\item $r = AAAAAAAAAA$.  $s$ covers $r$, since $s$ hits at every base.
\end{itemize}

@@

@@

>>>>

\begin{definition}
A \underline{repeat element family discriptor} (refd) is a sequence over
$\Sigma=\{A,C,G,T,*\}$ (where $*$ denotes unspecified bases).
\end{definition}

{\bf Note:} We will refer to sub-refds as consistent with the concept
of a substring, and denote then as consistent with Python.  (That is,
if $r =  AAAC*CGG$, then the sub-refd $r_{3:6}$, or $r[3:6]$, is the refd
$C*C$).  Generally, we will define all standard string operators on
refds unless otherwise noted.  (e.g. $\cdot$ will be the concatenation
operator, $|r|$ will denote total length, etc...)

@@

\begin{definition}
Let $r$ and $r'$ be two $l$-mers.
\begin{itemize}
\item $r$ and $r'$ are \underline{consistent} if,
  for any position $i$ in which they differ, one of them contains a
  $*$.  ($AAC*TT$ and $AA*GTT$ are consistent; $AAC*TT$ and $AAGGTT$
  are not.)
\item We will say $r$ \underline{generalizes} $r'$ if, wherever they differ,
  $r$ has a $*$.  Denote this as $r \vdash r'$.  ($AAC*TT \vdash AACGTT$, 
  $AAC*TT \not\vdash AA*CTT$.)
\item The \underline{merge} of $r$ and $r'$, denoted $r \oplus r'$, is
  the $l$-mer such that $s_i = r_i$ when $r_i=r'_i$,
  and $*$ otherwise. ($AAC*TT \oplus AA*GTT = AA**TT$.) 
\item The \underline{tight merge} of $r$ and $r'$, denoted $r \otimes r'$,
  is $r \oplus r'$ if $r$ and $r'$ are consistent, and undefined
  otherwise. ($AAC*TT \otimes AA*GTT = AA**GT$, but $AAC*TT \otimes AAAA*CGT$ is undefined.)
\item The \underline{compression} of consistent refds $r$ and $r'$, denoted $r\ominus r'$, 
  is the string $s$ where $s_i$ is: $*$ if both $r_i$ and $r'_i$ are
  $*$, other the non-$*$ character at $r_i$ or $r'_i$.  If $r$ is and
  $r'$ are not consistent, $r \ominus r'$ is undefined.  ($AA**C
  \ominus A**CC = AA*CC$.  But $AA**C \ominus AT**C$ is undefined
  (because of the disagreement in the second position).
\end{itemize}
\end{definition}

@@

\begin{definition}
Let $x$ and $y$ be REFDs. Define $x \circ y$ as follows.  Let $i$ be
the largest $i$ such that $x_{|x|-i:}$ and $y_{:i}$ are consistent, let
$b = x_{|x|-i:} \ominus y_{:i}$, let $a = x_{:|x|-i}$ and $c = y_{i:}$.  
$x \circ y = a \cdot b \cdot c$.
\end{definition}
{\bf Examples:}
\begin{itemize}
\item $AAACCC \circ C*CTTT = AAACCCTTT$.
\item $AAACCC \circ CTCTTT = AAATCCTCTTT$.
\item $AAACC* \circ CC*TTT = AAACC*TTT$.
\end{itemize}

@@

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the seeded
$l$-mer, denoted $\sigma_s(r)$, is created by removing all bases at
positions corresponding to 0s.
\end{definition}
Example: $\sigma_{11011}(AACGG) = AAGG$.

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
generlized $l$-mer (w.r.t. $s$), or $\gamma_s(r)$, is the refd created by replacing
all basses corresponding to 0s with * characters.
\end{definition}
Example: $\gamma_{110101}(AACGTG) = AA*G*G$.

@@

\begin{definition} 
A spaced seed $s$ of length $l$ \underline{covers} and refd $r$ is, for every $j$,
there is an $i \leq j$ such that $s$ is consistent with the sub-refd
$r_{i:i+|l|}$.
\end{definition}
{\bf Examples:} Let $s = 11011$.
\begin{itemize}
\item $r = AAAAA*AA$. $s$ covers $r$ as we can place $s$ at $i=0$ and
  $i=3$ and, between them, cover every character in $r$ without
  concflict between a 1 and a $*$.
\item $r = AAA*AA$.  $s$ does not cover $r$.  $s$ only hits at $i=1$,
  so there is no way to cover base 0.
\item $r = AAAAAAAAAA$.  $s$ covers $r$, since $s$ hits at every base.
\end{itemize}

@@

\subsection{REFD Decomposition}

Given an REDF $r$ and a seed $s$ of length $l$ ($l \leq |r|$), we want
to decompose $r$ into a set of overlapping $l$-mers.  These are the
$l$-mers that cannot be repeated outside of the repeat.

For example: if we have $G = \ldots AAAAACTTTTT\ldots AAAAAGTTTTT
\ldots$ (notice: 6 As), with the seed $1101$, we get the refd $AAAAAA*TT*TT$, which
deposes into $AA*AA$, $AA*TT$, and $TT*T$$.  If any string compatible
with any of these three occurs anywhere else, it breaks the family.
But if the string $AA*AC$ occurs somewhere else, it does not -- thats
irrelevant.  (In this example the minimal decomposition is unique -- I'm not
sure that this is always the case, or what multiple decompositions
mean.)

\begin{definition} 
Given a spaced seed $s$ of length $l$ and a an refd $r$ that is
covered by $s$, a \underline{decomposition} of $r$ w.r.t. $s$ is a set of coordinates $i$, $0 \leq i < |r|-l$, such
that (a) for each $i \in S$, $s$ hits $r$ at $i$, and (b) $r$ is
covered by $s$ using only those positions from $S$.
\end{definition}

Examples:
\begin{itemize}
\item $s=11011$, $r=AAAAA*CCCCC$, then the only decomposition is
  ${0,3,6}$.
\item $s=11011$, $r=AAAAAA*CCCCC$, then ${0,1,4,7}$ and ${0,4,7}$ are
  both decompositions.
\end{itemize}

\begin{definition}
A decomposition $S$ of $r$ with respect to $s$ is \underline{minimal}
if no element of $S$ can be removed without breaking the decomposition
properties. 
\end{definition}
In the above, the first set of the second example is not minimal.

@@

I want to get a definition of elementary repeats that paralleles the
one we had.  (And, hopefully, is provably equivilent for spaced seeds
without spaces.)  This is tricky: I'm not yet sure what the equivilent
is for the minimality condition (there is no subsequence of an
elementary repeat that occurs independently of the instances of the
elementary repeat).  Lets look why.

Suppose we have in the string out genome $AAAAATCCCCC$ and
$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
condition, and why they break:
\begin{enumerate}
\item No sub-refd can match any string in the genome outside of those
  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
  else wihtout breaking this family, since the second $l$-mer is
  irrelevant to construction.
\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
\end{enumerate}
